---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs


library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-1

# add code here
```

# Data 2

## Introduction and data

-   The data set, ufo_sightings.csv, is from the CORGIS Dataset project.

-   This data was originally collected by the National UFO Reporting Center (NUFORC) and tidied by a person named Sigmond Axel.

-   The observations collect data about UFO sightings worldwide, from date and time to city area, date documented, and more specific information about the description of the UFO, from size to lights.

## Research question

-   What is the most common location UFOs are sighted in? What is the most common size or shape of UFO? How can we analyze these reports to contribute to the dialogue about whether or not UFOs are evidence of alien life?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   https://www.rochester.edu/newscenter/nasa-brings-standards-of-evidence-to-the-search-for-ufos-527122/ Here is an article about how NASA will bring the scientific standard to the search for UFOs.

-   This article describes how NASA will be investigating UFOs, also referred to in the article as UAPs (unidentified aerial

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-2

# add code here
```

# Data 3

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

# add code here
```
