---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

Research Question #1 (dataset: ):

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-1

# add code here
```

# Data 2

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-2

# add code here
```

# Data 3

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

The source of the data is from the CORGIS (Collection Of Really Great, Interesting Situated) Dataset Project.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

coffee <- read_csv("data/coffee.csv")
glimpse(coffee)
```

INTRODUCTION AND DATA

Date and Method of Original Data Collection: These data were collected from the Coffee Quality Institute's review pages in January 2018. These data contain reviews of 1312 arabica and 28 robusta coffee beans from the Coffee Quality Institute's trained reviewers.

Each observation is a review on a type of coffee.

There doesn't seem to be any evident ethical concerns.

RESEARCH QUESTION

What is the relationship between the average altitude of the farm and the sweetness level of the coffee based on the country where the coffee was produced?

This question is important because the results of this study will demonstrate which country and at which altitude the coffee will be the sweetest or least sweet. This information will be important for coffee farmers who wish to produce sweet coffee (or vice versa) for their customers.

Our research will investigate the relationship between the average altitude of farm where the coffee is grown and the sweetness level of the coffee produced based on the country where the coffee was produced. We hypothesize that there will be a relationship between the altitude and sweetness as the altitude of where the coffee is grown can affect the type of environment that the coffee is grown in, which would have an impact on the taste of the coffee beans.

The variable Location.Country (country where the coffee is produced) is a categorical variable, whereas the variables Data.Scores.Sweetness (sweetness level of the coffee on a scale of 1-10) and Location.Altitude.Average (the average altitude of the farm in which the coffee was grown) are quantitative variables.

LITERATURE

This research article discusses the research into the sensory information about four different types of coffee, and how these attributes compared when coffee was grown at two different altitudes in the San Martin region of Peru. There was a 200 meter difference in the altitudes of coffee being grown and an ample amount -- 1500g -- of each type of coffee was grown in each location. Sensory factors "such as aroma, taste, aftertaste, acidity, body, consistency, balance, cleanliness of the cup, sweetness and beverage quality" were measured based on tables approved by the Specialty Coffee Association of America (SCAA). Moreover, the coffee tasters used to assess the coffee for this experiment were "certified by the Cooperativa Agraria Cafetalera y de Servicios Oro Verde". 

Our research builds on the research described in this article, because it is comparing sweetness (a sensory attribute) of coffee harvested at different altitudes in a variety of regions worldwide, not just the San Martin region of Peru.

\
