---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

```{r}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
```

# Introduction and Data

Our topic for this project is UFOs, also known as "Unidentified Flying Objects". UFOs as proof of alien life are largely written off as pseudoscience by the scientific community, but this does not take away from the fact that there are tens of thousands of sightings reported annually. In 2020, the Pentagon released videos of unidentified flying objects, sparking the intrigue of enthusiasts and skeptics alike. This is outlined in a New York Times article\*\* that also highlights how the government has been looking into unidentified aerial phenomena (UAP; another term for UFO) since the 1950s; articles like this one add merit to our research into consistencies in examining UFOs worldwide. Overall, our group chose to use this data set on UFOs as our research because it was a topic that really piqued our interest and that we knew would maintain its intrigue for the duration of the semester.

[Link to NYT article:]{.underline} https://www.nytimes.com/2020/04/28/us/pentagon-ufo-videos.html

**The research question we have landed on is the following:** Is the average length of a UFO encounter greater in the decades after 1970 when separated by country? Our hope is that this will help examine how time impacts UFO reports worldwide.

We chose the decade of 1970 as our indicator because it marked the release of "Close Encounters of the Third Kind," which could be a possible explanation in the case that our null hypothesis is rejected. This will be discussed deeper in the discussion section.

Ethical issues do not seem to be an issue with this data set, because we do not have any personal information about the sources of the data. All of the data is anonymous and has no identifying information which eliminates ethical concerns.

The data was uploaded by a man named Sam Arnold to Github on October 28, 2022. It originally came from the National UFO Reporting Center\* , and was tidied by a Github user named Sigmond Axel\*\*, originally in 2014. According to the link on Github, this data was initially uploaded as a part of a Tidy Tuesday project. The cases are over 80,000 UFO sightings worldwide.

[\*Link to NUFORC website:]{.underline} https://nuforc.org/

[\*\*Link to Github data site:]{.underline} https://github.com/planetsig/ufo-reports

There are three relevant variables we're examining from the data set: length of encounter, shape, and country.

-   Data.Encounter duration is measured in seconds and is classified as a "double" variable.

-   Year refers to the year the UFO was sighted in.

-   Finally, the Location.country variable reports what country the particular UFO was sighted in.

# Literature Review

In the late 1940s, the Air Force began a study called Project Sign (later called Project Grudge and then Project Blue Book) for the purpose of analyzing unidentified flying objects. Between 1948 and 1969 this project investigated over 12,000 sightings of unidentified aerial objects. At the end of the 22 years the study took data from (1947-1969), the Air Force filtered out aerial phenomena with explanations, ranging from "weather conditions" to balloons and satellites.

This study found that of the 12,618 UFO sightings over the period of the study, only 701 could not be explained. I could not find much about the design, but it seemed to be focused on dividing aerial phenomena into "unidentified" and "identified" objects.

This report on Project Sign connects to our research by demonstrating the necessity of expanding on visualizing what UFOs look like. It provides a good basis for initial research into UFOs, but our research will respond to a more detailed and recent data set. While this Air Force study was likely groundbreaking for its time, our research will work on evaluating similarities and differences in UFO sightings beyond just classifying a phenomenon as "unidentified". We also have fewer limitations on analyzing the data - in the context of the Air Force study, the motivation was evaluating the possibility of a national security threat. Our study provides a broader perspective, both beyond national security interests and beyond American borders.

In 1977, the Spielberg film "Close Encounters of the Third Kind"\* was released, and it became a blockbuster. The film tells the story of an ordinary American who becomes interested in UFOs after sighting one in his hometown. The popularity of this film led us to think it may have influenced people to "look" for UFOs or be more equipped to identify one in 1977 and after.

\*Link to info about the film: https://www.imdb.com/title/tt0075860/?ref\_=ttrel_ov_i

[**Link to study:**]{.underline} https://www.esd.whs.mil/Portals/54/Documents/FOID/Reading%20Room/UFOsandUAPs/2d_af_1.pdf

# Methodology

**Null hypothesis:** The mean encounter length between UFO encounters in the decades before 1977 is equal to encounter length in the decades after 1977.

**Alternative hypothesis:** The mean encounter length between UFO encounters in the decades before 1977 is less than in the decades after 1977.

```{r}
library(tidyverse)
library(tidymodels)
ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")

```

```{r}

#| label: hypothesis test

new_ufo_sightings <- ufo_sightings |>
  mutate(
    decade = case_when((str_detect(date_time, "194") == TRUE) ~ "1940s",
                       (str_detect(date_time, "195") == TRUE) ~ "1950s",
                       (str_detect(date_time, "196") == TRUE) ~ "1960s",
                       (str_detect(date_time, "197") == TRUE) ~ "1970s",
                       (str_detect(date_time, "198") == TRUE) ~ "1980s",
                       (str_detect(date_time, "199") == TRUE) ~ "1990s",
                       (str_detect(date_time, "200") == TRUE) ~ "2000s",
                       (str_detect(date_time, "201") == TRUE) ~ "2010s"
    )
  )

new_ufo_sightings <- new_ufo_sightings |>
  mutate(after_70 = case_when((str_detect(date_time, "194") == TRUE) ~ "NO",
                                  (str_detect(date_time, "195") == TRUE) ~ "NO",
                                  (str_detect(date_time, "196") == TRUE) ~ "NO",
                                  (str_detect(date_time, "197") == TRUE) ~ "NO",
                                  (str_detect(date_time, "198") == TRUE) ~ "YES",
                                  (str_detect(date_time, "199") == TRUE) ~ "YES",
                                  (str_detect(date_time, "200") == TRUE) ~ "YES",
                                  (str_detect(date_time, "201") == TRUE) ~ "YES"
  ))

ufo_us <- new_ufo_sightings |>
  filter(country == "us")

ufo_gb <- new_ufo_sightings |>
  filter(country == "gb")

ufo_ca <- new_ufo_sightings |>
  filter(country == "ca")

ufo_au <- new_ufo_sightings |>
  filter(country == "au")

ufo_de <- new_ufo_sightings |>
  filter(country == "de")

new_ufo_sightings |>
  drop_na()|>
  group_by(country, after_70) |>
  summarize(mean(encounter_length))

mean_us_no = 6968
mean_us_yes = 5450
mean_ca_no = 1242
mean_ca_yes = 30436
mean_gb_no = 900
mean_gb_yes = 9088

null_dist_us <- ufo_us |>
  filter(after_70 %in% c("NO", "YES")) |>
  specify(response = encounter_length, explanatory = after_70) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("NO", "YES"))

obs_stat_us = mean_us_no - mean_us_yes

null_dist_us |>
  get_p_value(obs_stat_us, direction = "two sided")

null_dist_ca <- ufo_ca |>
  filter(after_70 %in% c("NO", "YES")) |>
  specify(response = encounter_length, explanatory = after_70) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("NO", "YES"))

obs_stat_ca = mean_ca_no - mean_ca_yes

null_dist_ca |>
  get_p_value(obs_stat_ca, direction = "two sided")

null_dist_gb <- ufo_gb |>
  filter(after_70 %in% c("NO", "YES")) |>
  specify(response = encounter_length, explanatory = after_70) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("NO", "YES"))

obs_stat_gb = mean_gb_no - mean_gb_yes

null_dist_gb |>
  get_p_value(obs_stat_gb, direction = "two sided")
```

To answer our research question, we calculated the mean encounter length based on UFO shape. We did this to help condense over 80,000 observations into a format that wasn't overwhelming for a reader to understand. We then split up this mean time by country in order to compare the relationship in the context of our research question.

We removed Australia (AU) and Germany (country abbreviation DE) from our analysis because there were simply too few observations - for instance, Australia had no UFO sightings recorded prior to 1977. Our visualizations were made for mean encounter length across decades by country. We ran a hypothesis test for mean encounter length time of each country, got the p-value, and from this were able to reject the null hypothesis.

## Results

```{r}
# | label: graph

new_ufo_sightings |>
  drop_na() |>
  ggplot(aes(x = decade, y = encounter_length)) +
  geom_point() +
  facet_wrap(~country) +
  labs(x = "Decades", y = "Encounter Length (seconds)", title = "Relationship between the decade of UFO sighting and encounter time of each sighting")
```

In order to answer the question, which is to investigate what the relationship between the UFO shape and the length of the encounter time (in seconds) is, based on the country, we decided to form scatterplot graphs, based on countries, to visualize the relationship between UFO shape and mean length of time.

We have come to the conclusion that in general, there is no significant correlation between all the different UFO shapes and the mean encounter time. This is exhibited by the consistently low mean encounter times for almost all UFO shapes across countries. However, depending on the country, there are specific UFO shapes that result in high outlying mean encounter times. Specifically, for Great Britain (GB), the cone shaped UFOs yield, on average, a mean encounter time of 2000000 seconds. For Canada, the shapes categorized as "other" yields, on average, around 48000 seconds. This isn't as significant as that of the mean encounter time for cone shaped UFOs in GB, but it is also an outlier from the average encounter time of close to zero for all other UFO shapes.

Despite these significant outliers, the overwhelmingly low average of mean encounter times across UFO shapes and the countries examined demonstrate that there is not a significant relationship between the shape of a UFO and the length of time it was encountered on Earth.
